### 2018-10-02 paper.tex
Submitted

  We present Guild AI, an open source toolkit that facilitates model
  reuse for application development by applying traditional software
  packaging constructs to the domain of machine learning. Package
  management strategies are central to successful software
  ecosystems. Examples include npm for JavaScript \cite{npm}, APT for
  Debian \cite{apt}, and pip for Python \cite{pip}. Such tools excel
  at creating traditional software packages. Machine learning
  applications however present unique requirements that call for tool
  specialization. Guild AI combines the proven effectiveness of
  package managers with novel features to enable machine learning
  model reuse for application development across a variety of use
  cases.

@@

Packaged software is software that is organized in a way that can be
easily installed and used by \emph{users}. A user is a consumer of
software, rather than a producer or developer. A user may be software
developer, in the case of consuming software libraries or frameworks,
or she may be an end-user who uses software as a program with a
graphical or command line interface. In both cases, packaged software
may be provided in units that the user can install and start using
quickly and reliably.

@@

By running this command, a user is using a \emph{package manager}---in
this case the program \texttt|pip|---to install a package named
\verb|tensorflow|. With this innocuous command, the user implicitly expects
a number of things:

\begin{itemize}
\item The Python module \verb|tensorflow| should be installed.
\item Additional Python modules required by \verb|tensorflow| should
  also be installed.
\item Installed software should work as expected on the target system.
\item At a future date \verb|tensorflow| may be similarly uninstalled
  along with its no-longer-used required packages.
\item The act of installing and uninstalling \verb|tensorflow| should
  not break or otherwise destabilize other installed software on the
  target system.
\end{itemize}

Such package management facilities are essential to code reuse at the
system level. Package managers allow users to maintain complex
software systems with simple commands in a way that preserves their
integrity. They support rapid experimentation with new software as
users are free to install, use, and evaluate packages with the
confidence that they can safely remove that software later.

@@

A \emph{machine learning model} is challenging to define as models
have different contexts. In the most general context, models are
mathematical concepts or language that represents a system (ref). In
the context of machine learning, models may be represented by computer
source code, either explicitly as declared structures, or implicitly
by way of executable instructions that perform model related tasks. A
\emph{trained model} is a model that has learned some representation
of a system through an automated process of optimization.

@@

\subsubsection{Model zoos}

A common approach to sharing and reusing machine learning models is a
\emph{model zoo}---an online repository of reusable models that
includes source code, examples, documentation, and in some cases
pretrained models. The Caffe Model Zoo (ref) was one of the first
model zoos and has inspired several others (ref).

@@

\href{https://www.tensorflow.org/hub/}{TensorFlow Hub modules} and
\href{https://keras.io/applications/}{Keras Applications} are examples
of framework level facilities that provide a higher degree of model
reuse vis-\'a-vis model zoos. Both TensorFlow Hub modules and Keras
application support tools and patterns for packaging, distributing,
and using model source code and associated pretrained models. They
emphasize the use of \emph{transfer learning} for applying pretrained
models to data from different domains.

@@

A Guild AI \emph{package} is a Python wheel distribution (ref) that
was generated using the \verb|guild package| command. Guild AI
packages are generated using \verb|setuptools|. They contain Guild AI
specific metadata that advertises models they contain.

@@

A Guild AI \emph{resource} represents a set of source files that can
be made available as inputs to a model operation. Resources play an
important role in Guild AI's ability to effective use
models. Resources can be used to resolve project files, download files
from URLs, unpack archives, and reference files generated by other
operation.

@@

A \emph{run directory} is a directory used for a run. When an
operation is run, the run directory is used as the operating system
process working directory. Operations should write run artifacts
within the working directory to ensure they are associated with the
run.

By default, Guild AI generates a new directory for each run and
creates links to required sources within that directory. Once the
directory is prepared, Guild AI starts the run operating system
process, which may read sources within the run directory.

@@

Guild AI supports \emph{environments}, which are isolated runtime
environments used for installing and using models. Guild AI supports
environments created using
\href{https://virtualenv.pypa.io}{\emph{virtualenv}}.

@@

A user may run the \verb|guild help| command to show details about a
model, its operations, operation flags, and resource
dependencies. This is similar to the \emph{man} facility (ref) on many
systems.

@@

Guild AI also provides various methods for visualizing runs. Guild
View is an application provided with Guild AI that displays run
details including run files. Guild AI integrates TensorBoard (ref) as
a visualization tool for viewing TensorFlow event logs generated by
model operations.\footnote{Not all operations generate TensorFlow
  event logs---these must be explicitly written during a run by the
  operation.}

@@

Runs may also be \emph{exported} to a location using the
\verb|guild export| command. Exported runs are copied---or moved, if
specified---to a specified file system directory. This can be used to
create backups or to transfer runs to another location.

@@

When running Guild AI operations in an activated environment, the
following are environment specific:

\begin{itemize}
\item Installed packages and therefore installed models
\item Runs
\end{itemize}

@@

One or more runs may be deleted using the \verb|guild runs rm|
command. Runs may be deleted using status, operation, or label.

@@

Operation \emph{flags} specified in the form \verb|NAME=VALUE| as
arguments to the \verb|guild run| command are passed to the Python
main module as command line arguments in the form
\verb|--NAME VALUE|. For example, assuming an operation \verb|train|
with a main module \verb|train|, the command
\verb|guild run train epochs=100| would be translated to the Python
command \verb|python -m train --epochs 100|. In this way developers
can parameterize scripts using operation flags.\footnote{It is
  standard practice to use Python's argparse module to support named
  command line options following the getopt convention (ref).}

@@

The second method is to wrap upstream scripts and monkey patch (ref)
the Python environment as needed to change behavior. An example of
this can be found in the \verb|gpkg.slim| package, which patches the
TensorFlow \verb|slim.losses.softmax_cross_entropy| module to add
support for balanced weight training. This technique successfully adds
a useful feature while avoiding the need to modify the upstream train
support in TF Slim.\footnote{While one could argue that it would be
  better to submit a pull request with the new functionality to
  upstream authors, there are always cases where this is not an
  option. Patching in these cases may be the only way to fix issues or
  add features.}

@@

\begin{figure}
\begin{lstlisting}[
    caption=Quickly enabling model deploy,
    label={lst:deploy}]
- package: classifiers
  requires: [server-utils]

@@

Guild AI supports YAML (ref) as its configuration language for both
Guild files and user configuration. It uses the PyYAML Python library
(ref) for low level parsing.

@@

Guild AI is freely available under the Apache 2.0 open source
license. It has been under active development for the last two years
and undergone significant revision and updates during that time based
on user feedback and real world application development.

There have been several packaging initiatives that have produced
reusable models (ref) that fulfill the goals outlined in this
paper. Specifically, the packages support:

>>>>

  We present Guild AI, an open source toolkit that facilitates model
  reuse for application development by applying traditional software
  packaging constructs to the domain of machine learning. Package
  management strategies are central to successful software
  ecosystems. Examples include npm for JavaScript \cite{npm}, APT for
  Debian \cite{apt}, and pip for Python \cite{pip}. Such tools excel
  at creating traditional software packages. Machine learning
  applications, however, present unique requirements that call for
  tool specialization. Guild AI combines the proven effectiveness of
  package managers with novel features to enable machine learning
  model reuse for application development across a variety of use
  cases.

@@

Packaged software is software that is organized in a way that can be
easily installed and used by \emph{users}. A user is a consumer of
software, rather than a producer or developer. A user may be a
software developer, in the case of consuming software libraries or
frameworks, or she may be an end-user who uses software as a program
with a graphical or command line interface. In both cases, packaged
software may be provided in units that the user can install and begin
using quickly and reliably.

@@

By running this command, a user is using a \emph{package manager}---in
this case the program \verb|pip|---to install a package named
\verb|tensorflow|. With this innocuous command, the user implicitly expects
a number of things:

\begin{itemize}
\item The Python module \verb|tensorflow| should be installed.
\item Additional Python modules required by \verb|tensorflow| should
  also be installed.
\item Installed software will work as expected on the target system.
\item At a future date \verb|tensorflow| may be similarly uninstalled
  along with its no-longer-used required packages.
\item The act of installing and uninstalling \verb|tensorflow| should
  not break or otherwise destabilize other installed software on the
  system.
\end{itemize}

Such package management facilities are essential to code reuse at the
system level. Package managers allow users to maintain complex
software systems with simple commands in a way that preserves their
integrity. They support rapid experimentation with new software as
users are free to install and evaluate packages with the confidence
that they can safely remove that software later.

@@

A \emph{machine learning model} is challenging to define as models
have different contexts. In the most general context, models are
mathematical concepts or language that represents a system. In the
context of machine learning, models may be represented by source code,
either explicitly as declared structures, or implicitly by way of
executable instructions that perform model related tasks. A
\emph{trained model} is a model that has learned some representation
of a system through an automated process of optimization.

@@

\subsubsection{Model Zoos}

A common approach to sharing and reusing machine learning models is a
\emph{model zoo}---an online repository of reusable models that
includes source code, examples, documentation, and in some cases
pretrained models. The Caffe Model Zoo \cite{caffezoo} was one of the
first model zoos and has inspired several others \cite{tfzoo,mxzoo}.

@@

TensorFlow Hub modules \cite{tfhub} and Keras Applications
\cite{kerasapps} are examples of framework level facilities that
provide a higher degree of model reuse vis-\'a-vis model zoos. Both
TensorFlow Hub modules and Keras application support tools and
patterns for packaging, distributing, and using model source code and
associated pretrained models. They emphasize the use of \emph{transfer
  learning} for applying pretrained models to data from different
domains.

@@

A Guild AI \emph{package} is a Python wheel distribution \cite{wheels}
that was generated using the \verb|guild package| command. They
contain Guild AI specific metadata that advertises models they
contain.

@@

A Guild AI \emph{resource} represents a set of source files that can
be made available as inputs to a model operation. Resources play an
important role in Guild AI's ability to effectively use
models. Resources can be used to resolve project files, download files
from URLs, unpack archives, and reference files generated by other
operation.

@@

A \emph{run directory} is a directory used for a run. When an
operation is run, the run directory is used as the process working
directory. Operations should write run artifacts within the working
directory to ensure they are associated with the run.

By default, Guild AI generates a new directory for each run and
creates links to required sources within that directory. Once the
directory is prepared with required sources, Guild AI starts the run
process, which may read sources within the run directory.

@@

Guild AI supports \emph{environments}, which are isolated runtime
environments used for installing and using models. Guild AI supports
environments created using {\emph{virtualenv} \cite{venv}.

@@

A user may run the \verb|guild help| command to show details about a
model, its operations, operation flags, and resource
dependencies. This is similar to the \emph{man} facility on many
systems.

@@

Guild AI also provides various methods for visualizing runs. Guild
View is an application provided with Guild AI that displays run
details including run files. Guild AI integrates TensorBoard \cite{tb}
as a visualization tool for viewing TensorFlow event logs generated by
model operations.\footnote{Not all operations generate TensorFlow
  event logs---these must be explicitly written during a run by the
  operation.}

@@

Runs may also be \emph{exported} to a location using the
\verb|guild export| command. Exported runs are copied---or moved, if
specified---to a different file system directory. This can be used to
create backups or to transfer runs to another location.

@@

When running Guild AI commands in an activated environment, the
following are environment specific:

\begin{itemize}
\item Installed packages and models
\item Runs
\end{itemize}

@@

One or more runs may be deleted using the \verb|guild runs rm|
command. Runs may be deleted by status, operation, or label.

@@

Operation \emph{flags} specified in the form \verb|NAME=VALUE| as
arguments to the \verb|guild run| command are passed to the Python
main module as command line arguments in the form
\verb|--NAME VALUE|. For example, assuming an operation \verb|train|
with a main module \verb|train|, the command
\verb|guild run train epochs=100| would be translated to the Python
command \verb|python -m train --epochs 100|. In this way developers
can parameterize scripts using operation flags.\footnote{It is
  standard practice to use Python's argparse module to support named
  command line options following the getopt convention \cite{getopt}.}

@@

The second method is to wrap upstream scripts and monkey patch the
Python environment as needed to change behavior. An example of this
can be found in the \verb|gpkg.slim| package, which patches the
TensorFlow \verb|slim.losses.softmax_cross_entropy| module to add
support for balanced weight training. This technique successfully adds
a useful feature while avoiding the need to modify the upstream train
support in TF Slim.\footnote{While one could argue that it would be
  better to submit a pull request with the new functionality to
  upstream authors, there are always cases where this is not an
  option. Patching in these cases may be the only way to fix issues or
  add features.}

@@

\begin{figure}
\begin{lstlisting}[
    caption=Enabling model deploy,
    label={lst:deploy}]
- package: classifiers
  requires: [server-utils]

@@

Guild AI supports YAML \cite{yaml} as its configuration language for both
Guild files and user configuration. It uses the PyYAML Python library
(ref) for low level parsing.

@@

Guild AI is freely available under the Apache 2.0 \cite{guild} open source
license. It has been under active development for the last two years
and undergone significant revision and updates during that time based
on user feedback and real world application development.

There have been several packaging initiatives that have produced
reusable models \cite{pkg} that fulfill the goals outlined in this
paper:
